MS Prelim Data K23, using Depression T=3.09 mask

### 2022-07-01 msdepression/scripts/K23_prelim_data_using_dep_T_3_09_mask_20220630.Rmd
Add files via upload

---
title: "MS Prelim Data K23"
author: "Erica Baller"
date: "2/25/2022"
output:
  html_document: default
  pdf_document: default
---

@@

Ideas:

Overall:

[] demographics - tables made below

[] imaging parameters

Aim 1: lesion network mapping
[] some picture showing the # of people who are scanned and the # of people with depression. Maybe an adaptation of Syndey's picture somehow. 
[] mimosa picture
[] picture of filtering pipeline
[] comparison of some of the histograms for high volume loss similar to prev literature(CN II, slf), not so much in other areas
[] results of age analysis [maybe/maybe not, to prove the method works]?
[] results of disease in vs outside of depression network (corrected), lead with this?
[] results total proportion lost (single p)
[] results of dep vs controls across all regions,(uncorrected)
[] results of dep vs controls in depression network (uncorrected)

[] prelim network enrichment via spin-testing or just citation?
- do I want to validate here or somewhere else with neurosynth maps?

Aim 2: prospective
[] power analysis
[] CNB (Bart has great Figure 3)
[] Goassess

Aim 3
[] Hydra picture
[] hydra results from my paper
[] acquisition protocol (T1, t2, flair, dwi)

@@

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller. It does however highlight the

@@

@@

Depression network:
 [1] "AF_L"    "AF_R"    "C_PHP_R" "EMC_L"   "EMC_R"   "FAT_L"   "FAT_R"   "IFOF_L"  "IFOF_R"  "ILF_R"   "MdLF_L" 
[12] "MdLF_R"  "PAT_L"   "PAT_R"   "SLF2_L"  "SLF2_R"  "SLF3_L"  "SLF3_R"  "V"       "AR_R"    "CBT_L"   "CBT_R"  
[23] "CPT_F_L" "CPT_F_R" "CPT_O_L" "CPT_O_R" "CPT_P_L" "CPT_P_R" "CS_A_L"  "CS_A_R"  "CS_P_L"  "CS_P_R"  "CS_S_L" 
[34] "CS_S_R"  "CST_L"   "CST_R"   "DRTT_R"  "ML_L"    "RST_L"   "RST_R"   "TR_A_L"  "TR_A_R"  "TR_P_L"  "TR_P_R" 
[45] "TR_S_L"  "TR_S_R"  "CC"

Depression 10% network
 [1] "AF_L"    "AF_R"    "C_PHP_R" "FAT_L"   "FAT_R"   "MdLF_L"  "MdLF_R"  "SLF2_L"  "SLF2_R"  "SLF3_R"  "CBT_L"  
[12] "CBT_R"   "CPT_O_R" "CS_P_R"  "TR_P_R"

@@

    *Sex effects:
      - In the final good MIMoSA group, there were 64 unique males and 316 unique females, 83% women
      - Individual fascicle results underwhelming, only 3 regions meet P<0.05 (unc), SCP, CPT_O_R, OR_R
      - no association or within dep network findings
      - Sex effects interaction model (fascicle ~ depGroupVar*osex + PAT_AGE_AT_EXAM)
        - Dep effects - same as above, 12 fascicles (p<0.05 unc)
        - Sex effects - same as above, with the edition of CNVIII_R (p<0.05 unc)
        - age*sex interactions - SLF1_L and CNVIII_R (p<0.05 unc)
      - Analyses to consider at a later date
        [] some sort of matching

    *Depression over all fascicles:
      - The exciting news of the day! With our new depression group (moving the meds people), we have findings
      - total proportion lost - p = 0.049, depressed (27%) more loss than healthy(24%)
      - total proportion lost within depression network (47 total) - trend (0.059) between depressed (35%) and healthy(30%)
      - total proportion lost within depression 10% network (15 fascicles total) - p = 0.049 (depressed (33%) have more loss than healthy (27%))
      - total proportion lost within depression 5% network (24 fascicles total) -  p = 0.056 (depressed (32%) have more loss than healthy (27%))
      - total proportion lost outside depression network (40 fascicles total)- p = 0.037 (depressed (18%) more than healthy(16%))
      - take home - people with a lifetime history of depression have more disease than nondepressed individuals. This finding is not specific to the depression network, but is overall.

    *MS disease and depression network overlap:
      - Ms disease within depression network (35%) versus outside (17%), p < 2.2* 10^-16
      - MS disease within depression 10% network (31%) versus outside (17%), p < 2.2 * 10^-16
      - MS disease within depression 5% network (30%) versus outside (17%), p < 2.2 * 10^-16
      - Linear Regressions (essentially matching t tests above): 
         +depressed v healthy within dep network in (p = 0.061); 
         +dep vs healthy within dep 10% (p = 0.049);
         +dep vs healthy within non dep network (p = 0.037)
      - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis. Suggests a final possible common pathway? Shared biology? This may also in part explain why rates of depression in MS are much higher than rates of depression in other autoimmune conditions. 
    *Depression by individual fascicle:
      - Nothing corrects at this time, unfortunately. 
      - P<0.05, uncorrected findings (* indicates association network):
          -AF_R	0.01168289	*		
          -FAT_R	0.01828607 *			
          -SLF2_R	0.01901378 *		
          -VOF_R	0.04741569 *			
          -CPT_F_R	0.02887163 			
          -CPT_O_L	0.02238344			
          -CPT_P_R	0.02053104			
          -CS_S_R	0.03783654			
          -DRTT_R	0.00695066			
          -ML_R	0.02511932	
    Fascicle loss x PHQ9:
     - Nothing significant by individual fascicle, within association network, by fascicle within the depression network
     - similarly, no significant findings looks at means of all fascicle loss, within dep network, 10% and 5% and non-depression network

@@

write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

########
#fascicle volumes in depression network
########
fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), header = T, sep = ",")

@@

write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

@@

#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all.csv", row.names = F, quote = F, col.names = F)

fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)

@@

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
fascicle_bundle_mapping <- get_fascicle_bundle_mapping()

@@

######## Histograms for depression network
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20)

@@

dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
summary(dep_vs_healthy) #p=0.494, cohen d = 0.2

@@

#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049

@@

#look at within depression network specifically - trend, p = 0.059
total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 36%, mean non_dep = 32%, p = 0.071

#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
#yes, p = 0.048
total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048

#look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
#yes, 
total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.056

#look at outside depression network, yes p = 0.037
total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.037

### Age actually makes things worse
#dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
#summary(dep_vs_healthy_w_age)

@@

#test whether there is differentially more volume loss in depression versus nondepression network
volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 35%, mean non_dep_net =  17%, p 2.2 x 10^-16

volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 31%, mean non_dep_net = 17%, p 2.2 * 10^-16

volume_lost_dep5_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent,
dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 17%, p 2.2 * 10^-16

@@

loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #non sig, p = 0.061
loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.0492
loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.037
```

@@

#visreg
#sapply(fascicle_sex_lm,visreg)
```

@@

>>>>

---
title: "MS Prelim Data K23, using Depression T=3.09 mask"
author: "Erica Baller"
date: "6/30/22"
output:
  html_document: default
  pdf_document: default
---

@@

@@

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller.

@@

Lastly, unlike the prelim K data analyses, this uses a depression mask that I generated which is slightly more leniant than the one I used for my first iteration (Matt thresholded that at 3.49. I threshold this one a 3.09).The only change I made it changing which file is fed into the depression analysis.

@@

Depression network (* where there are new regions as compared to the 3.49 mask, now we are up to 62 regions):
[1] "AF_L"   "AF_R"    "C_FP_L"*  "C_FP_R"*  "C_PHP_L"* "C_PHP_R" "C_R_R"*   "EMC_L"   "EMC_R"   "FAT_L"   "FAT_R" [12] "IFOF_L"  "IFOF_R"  "ILF_L"*   "ILF_R"   "MdLF_L"  "MdLF_R"  "PAT_L" "PAT_R"   "SLF1_L"*  "SLF1_R"*  "SLF2_L" [23] "SLF2_R"  "SLF3_L"  "SLF3_R"  "VOF_R"*   "CB_L"*    "CB_R"*    "MCP"*     "V"   "AR_L"*    "AR_R"    "CBT_L" [34] "CBT_R"   "CPT_F_L" "CPT_F_R" "CPT_O_L" "CPT_O_R" "CPT_P_L" "CPT_P_R" "CS_A_L"  "CS_A_R"  "CS_P_L"  "CS_P_R" 
[45] "CS_S_L"  "CS_S_R"  "CST_L"   "CST_R"   "DRTT_L"*  "DRTT_R"  "F_R"*   "ML_L"    "RST_L"   "RST_R"   "TR_A_L" 
[56] "TR_A_R"  "TR_P_L"  "TR_P_R"  "TR_S_L"  "TR_S_R"  "AC"*     "CC"

Depression 10% network (stars where there are new regions as compared to the 3.49 mask, 21 regions)
 [1] "AF_L" "AF_R"    "C_PHP_R" "EMC_R"*   "FAT_L"*   "FAT_R"   "MdLF_L"  "MdLF_R"  "PAT_L"*   "PAT_R"*   "SLF2_L"
[12] "SLF2_R"  "SLF3_R"  "CBT_L"   "CBT_R"   "CPT_O_R" "CS_P_R"  "TR_P_L"*  "TR_P_R"  "TR_S_L"*  "TR_S_R" *

@@

    *Sex effects:
      - In the final good MIMoSA group, there were 64 unique males and 316 unique females, 83% women
      - Individual fascicle results underwhelming, only 3 regions meet P<0.05 (unc), SCP, CPT_O_R, OR_R
      - no association or within dep network findings
      - Sex effects interaction model (fascicle ~ depGroupVar*osex + PAT_AGE_AT_EXAM)
        - age*sex interactions - SLF1_L and CNVIII_R (p<0.05 unc)
      - Analyses to consider at a later date
        [] some sort of matching

    *Depression over all fascicles (also, for within/outside dep network, used rowMeans and rowSums, doesn't change anything w/r/t statistics, you just get sums instead of means for means between groups):
      - The exciting news of the day! With our new depression group (moving the meds people), we have findings
      - total proportion lost - p = 0.049, depressed (27%) more loss than healthy(24%)
      - total proportion lost within depression network (xx total) - trend (0.054) between depressed (33%) and healthy(28%)
      - total proportion lost within depression 10% network (xx fascicles total) - p = 0.055 (depressed (32%) have more loss than healthy (27%))
      - total proportion lost within depression 5% network (xx fascicles total) -  p = 0.053 (depressed (32%) have more loss than healthy (27%))
      - total proportion lost outside depression network (xx fascicles total)- p = 0.042 (depressed (14%) more than healthy(13%))
      - take home - people with a lifetime history of depression have more disease than nondepressed individuals. This finding is not specific to the depression network, but is overall.

    *MS disease and depression network overlap:
      - Ms disease within depression network (31%) versus outside (14%), p < 2.2* 10^-16
      - MS disease within depression 10% network (30%) versus outside (14%), p < 2.2 * 10^-16
      - MS disease within depression 5% network (30%) versus outside (14%), p < 2.2 * 10^-16
      - Linear Regressions (essentially matching t tests above): 
         +depressed v healthy within dep network in (p = 0.055); 
         +dep vs healthy within dep 10% (p = 0.053);
         +dep vs healthy within non dep network (p = 0.042)
      - Take home: Multiple sclerosis white matter lesions co-locate with brain regions previously associated with depression, irrespective of diagnosis. Suggests a final possible common pathway? Shared biology? This may also in part explain why rates of depression in MS are much higher than rates of depression in other autoimmune conditions. 
    *Depression by individual fascicle:
      - Nothing corrects at this time, unfortunately. 
      - P<0.05, uncorrected findings:
          -AF_R	0.03235420			
          -FAT_R	0.03337075			
          -SLF2_R	0.04066933			
          -CPT_F_R	0.04932598			
          -CPT_O_L	0.04346334			
          -CPT_P_R	0.03471933			
          -DRTT_R	0.01914199			
          -ML_R	0.01259474			
          -TR_S_R	0.01724321	
    Fascicle loss x PHQ9:
     - Nothing significant by individual fascicle, within association network, by fascicle within the depression network
     - similarly, no significant findings looks at means of all fascicle loss, within dep network, 10% and 5% and non-depression network

     Depression by sex: there is an ineraction, interestingly:
     - Males with depression have higher loss in depression network, and males without depression have less loss in depression network, as compared to females. I.e. females, depression yes/no less related to lesions in/out dep network

     * Comparison of proportion overlap between depression network 3.49, 3.09

     * Proportion lost per volume x proportion overlap in depression network (prop lost * dep proportion)
     - 3.09;  only uncorrected values when looking at the relationship between depression diagnosis and lesion hit (can only look at within depression map in this analysis)
     -  AF_R_3_09	0.012274944
        FAT_R_3_09	0.022765660			
        SLF2_R_3_09	0.021441998			
        VOF_R_3_09	0.049555998			
        CPT_F_R_3_09	0.031783931			
        CPT_O_L_3_09	0.023932661			
        CPT_P_R_3_09	0.020599230			
        CS_S_R_3_09	0.042603557			
        DRTT_R_3_09	0.008120982			
        TR_A_R_3_09	0.049682290

@@

#write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

########
#fascicle volumes in depression network
########

#### THIS WAS CHANGED 6/30 from streamline_volume_within_dep_network.csv to streamline_volume_within_dep_network_3_09.csv

fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), header = T, sep = ",")

@@

#changed this 6/30 by adding 3_09, need this for bundle mapping
fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")
write.table(fascicle_volumes_dep_network, file=fascicle_names_dep_network_grouping, sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

@@

#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all_3_09.csv", row.names = F, quote = F, col.names = F)

fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent_3_09.csv", row.names = F, quote = F, col.names = F)

@@

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)

@@

######## Histograms for depression network
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Volume per Fascicle Affected", breaks = 20)

@@

dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
summary(dep_vs_healthy) #p=0.0494

@@

#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049, cohen d = 0.17

@@

#look at within depression network specifically - trend, p = 0.054
total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 28%, p = 0.054

#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
#yes, p = 0.055
total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.055

#look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
#yes, p=0.053
total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.053

#look at outside depression network, yes p = 0.042
total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 14%, mean non_dep = 13%, p = 0.042

### Age actually makes things worse, switched age and dep group var around, values are the same
#however, when I send an anova over the linear model, depression group var becomes significant
dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ PAT_AGE_AT_EXAM + depGroupVar , data = dep_and_healthy_groups_for_ICD_analysis)
summary(dep_vs_healthy_w_age) #dep group var p = 0.101
dep_vs_healthy_w_age_anova=anova(dep_vs_healthy_w_age) # p = 0.044

```{r depression_unique_all_fascicles_together_in_dep_networks_sum_rather_than_mean}

#### SUM DOESN"T CHANGE STATS AT ALL
#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowSums(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowSums(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowSums(across(fascicle_names_dep_network_10_percent))) %>%
   mutate(total_fascicle_prop_lost_indepnet_5_percent =
           rowSums(across(fascicle_names_dep_network_5_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowSums(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
summary(dep_vs_healthy) #p=0.0494

#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049, cohen d = 0.17

#look at within depression network specifically - trend, p = 0.054
total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 20.2, mean non_dep = 17.4, p = 0.054

#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
#yes, p = 0.055
total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 6.7, mean non_dep = 5.6, p = 0.055

#look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
#yes, p=0.053
total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 10.2, mean non_dep = 8.6, p = 0.053

#look at outside depression network, yes p = 0.042
total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 3.6, mean non_dep = 3.2, p = 0.042

### Age actually makes things worse, switched age and dep group var around, values are the same
#however, when I send an anova over the linear model, depression group var becomes significant
dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ PAT_AGE_AT_EXAM + depGroupVar , data = dep_and_healthy_groups_for_ICD_analysis)
summary(dep_vs_healthy_w_age) #dep group var p = 0.101
dep_vs_healthy_w_age_anova=anova(dep_vs_healthy_w_age) # p = 0.044

@@

#test whether there is differentially more volume loss in depression versus nondepression network
volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 31%, mean non_dep_net =  14%, p 2.2 x 10^-16

volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 14%, p 2.2 * 10^-16

volume_lost_dep5_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent,
dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 14%, p 2.2 * 10^-16

@@

loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #non sig, p = 0.055
loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.056
loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.042

@@

#uncorrected
fascicle_anova_mm_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
print(fascicle_anova_mm_p05_unc)
#visreg
#sapply(fascicle_sex_lm,visreg)
```

@@

```
```{r compare_proportion_overlap_in_dep_3_49_and_3_09_maps}
fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

dep_map_3_49 <-
  read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), sep = ",", header = T)
dep_map_3_09 <-
  read.csv(paste0(homedir, "results/streamline_volume_within_dep_network_3_09.csv"), sep = ",", header = T)

dep_map_together <- data.frame(dep_map_3_49$fascicle, round(dep_map_3_49$prop_in_mask,3), round(dep_map_3_09$prop_in_mask,3))
names(dep_map_together) <- c("fascicle", "dep_map_3_49_prop", "dep_map_3_09_prop")

#get fascicle names vectors
#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
names_dep_map_3_49 <- gsub("^(.*)", "\\1_3_49", fascicle_names)
names_dep_map_3_09 <- gsub("^(.*)", "\\1_3_09", fascicle_names)

#multiply each row (Margin=2 equals by row), each %overlap w/depression map of that fascicle, this is like a dysconnectome score, but doesn't take into account the STRENGTH(T) of the voxels in the depression mask. Using the third column to the end of the fascicle_proportions data_frame because the first 2 are empi and date
lesion_prop_x_prop_in_dep_mask_3_49<-sweep(fascicle_proportions[,3:(dim(fascicle_proportions)[2])], MARGIN=2, dep_map_together$dep_map_3_49_prop, `*`)
names(lesion_prop_x_prop_in_dep_mask_3_49) <- names_dep_map_3_49

lesion_prop_x_prop_in_dep_mask_3_09<-sweep(fascicle_proportions[,3:(dim(fascicle_proportions)[2])], MARGIN=2, dep_map_together$dep_map_3_09_prop, `*`)
names(lesion_prop_x_prop_in_dep_mask_3_09) <- names_dep_map_3_09

#merge them all together, big data frame, the proportion of each fascicle that is affected per person, and the proportion of that fascicle with the proportion of depression overlap for that fascicle

fascicle_volumes_dep_w_lesion_prop_x_prop_in_dep_mask <- data.frame(fascicle_proportions, lesion_prop_x_prop_in_dep_mask_3_49, lesion_prop_x_prop_in_dep_mask_3_09)

df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_volumes_dep_w_lesion_prop_x_prop_in_dep_mask, by = c("EMPI", "EXAM_DATE")) #n = 2962

#get unique first scan, n = 380

dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>%
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

### for 3_09

#### map the fiber bundles
fascicle_names_dep_network_grouping <- paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping_3_09.csv")

fascicle_bundle_mapping <- get_fascicle_bundle_mapping_generic_dep_mask(dep_network_grouping_file = fascicle_names_dep_network_grouping)

#0 is not in, 1 is between 0 and 10% mask, 2 is 10% mask, >0 is all
names_dep_map_3_09_excluding_na_cols <- names_dep_map_3_09[which(fascicle_bundle_mapping$inDepNetwork_vector > 0)]

#lm
